---
title: "Clustering tests"
author: "Sebastien Abilla"
date: "31/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```
## libraries and data
```{r libraries and data}
library(dplyr);library(readr);library(fastcluster)
library(cluster);library(factoextra);library(magrittr)
data <- read_csv("Data/Data.csv")
```


## hclust algorithm
Proposed method, I used <https://www.stat.berkeley.edu/~s133/Cluster2a.html> to understand the basics.

```{r hclust}
dist <- round(dist(data))
method = c("complete","single","average","centroid","median","ward","mcquitty")
hclusttable_groups3 <- tibble(observation = 1:27)
par(mfrow=c(2,2))
for (i in 1:7) {
  cluster <- hclust(dist, method = method[i]) 
  plot(cluster)
  groups3 <- cutree(cluster,3) 
  hclusttable_groups3 <- tibble(hclusttable_groups3, groups3 = groups3)
  names(hclusttable_groups3)[i+1] <- paste("hc",method[i])
}
```

## k-means algorithm
Maybe the most used method nowadays. "Powerful" computers required. (Powerful... compared to 30 years before)

#### Evaluate the number of clusters

```{r}
method = c("wss", "silhouette", "gap_stat")
a <- fviz_nbclust(data, kmeans, method = method[1],k.max = 20)+
  labs(subtitle = method[1])
b <- fviz_nbclust(data, kmeans, method = method[2],k.max = 20)+
  labs(subtitle = method[2])
c <- fviz_nbclust(data, kmeans, method = method[3],k.max = 20)+
  labs(subtitle = method[3])
gridExtra::grid.arrange(a,b,c, ncol=2)
```

1 or 2 clusters are recommanded, but lets do 3. :) 

#### Examples

k-means use randomness so it didn't always give the same answer.

```{r}
a <- fviz_cluster(kmeans(data, 3), data = data, frame.type = "convex")
b <- fviz_cluster(kmeans(data, 3), data = data, frame.type = "convex")
c <- fviz_cluster(kmeans(data, 3), data = data, frame.type = "convex")
d <- fviz_cluster(kmeans(data, 3), data = data, frame.type = "convex")
gridExtra::grid.arrange(a,b,c,d)
```

So, let's do it several times.

#### re-run the function

```{r}
kmeans3 <- replicate(20, {
km <- kmeans(data, 3)
tibble(clu = km$cluster) %>% 
  group_by(clu) %>% mutate(n = n()) %>% ungroup() %>%
  mutate(new_group = ifelse(n == max(n), 1, ifelse(n == min(n), 3, 2))) %>%
  pull(new_group) %>% unlist
})
meankmeans <- round(rowMeans(kmeans3))
```

I select the cluster based on mean. A selection based on max frequency may be better, but it's not obvious for me how to code it. I will have to think/search how to do it.

## Conclusion : All methods compared

```{r}
clusters <- tibble(hclusttable_groups3, kmeans = meankmeans)
knitr::kable(clusters)
```

As suggested by the k-means evaluation, there is probably not enough data yet to really make 3 clusters. There are 2 groups and one "strange" number 5.